---
title: "easier data"
author: "Oscar Lapuente-Santana"
package: easierData
output: 
    rmarkdown::html_document
vignette: >
    %\VignetteIndexEntry{easier data}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load easier data

The `easierData` data package includes an exemplary dataset to showcase the `easier` package:

* **Mariathasan2018_PDL1_treatment**: exemplary bladder cancer dataset with samples from 192 patients. The processed data is publicly available from Mariathasan et al. "TGF-B attenuates tumour response to PD-L1 blockade by contributing to exclusion of T cells", published in Nature, 2018 [doi:10.1038/nature25501](https://doi.org/10.1038/nature25501) via [`IMvigor210CoreBiologies`](http://research-pub.gene.com/IMvigor210CoreBiologies/) package under the CC-BY license.
  
The `easierData` data package also includes multiple data objects so-called internal data of `easier` package since they are indispensable for the functional performance of the package. This includes:

- **opt_models**: cancer-specifc model feature parameters learned, for each quantitative descriptor during training using multi-task learning with randomized cross-validation repeated 100 times. For each quantitative descriptor, 1000 models are available (100 per task).
    
- **opt_xtrain_stats**: cancer-specific features mean and standard deviation of the training set used during
      randomized cross-validation repeated 100 times. It is required for normalization of the test set.
    
- **TCGA_mean_pancancer**: mean of the TPM expression of each gene across all TCGA cancer types. 
      It is required for normalization of input TPM gene expression data.
    
- **TCGA_sd_pancancer**: standard deviation of the TPM expression of each gene across all TCGA cancer types. 
      It is required for normalization of input TPM gene expression data.
    
- **cor_scores_genes**: genes used to define correlated scores of immune response.
    
- **intercell_networks**: cancer-specific intercellular networks.
    
- **lr_frequency_TCGA**: frequency of each ligand-receptor pair feature across the whole TCGA database.
    
- **group_lrpairs**: information to group ligand-receptor pairs because of sharing the same gene", either as ligand or receptor.
    
- **HGNC_annotation**: gene symbols approved annotations.
    
- **scores_signature_genes**: gene signatures for each score of immune response: CYT, TLS, IFNy, Ayers_expIS, Tcell_inflamed, Roh_IS, Davoli_IS, chemokines, IMPRES, MSI and RIR.

The contents of the package can be seen by querying annotationHub for the package name:
```{r}
suppressPackageStartupMessages({
    library(ExperimentHub)
    library(easierData)
})


eh <- ExperimentHub()
query(eh, "easierData")
```

The individual data objects can be accessed using either their ExperimentHub
accession number, or the convenience functions provided in this package. For
instance to access the Mariathasan2018_PDL1_treatment example dataset:

```{r, message=FALSE}
mariathasan_dataset <- eh[["EH6677"]]
mariathasan_dataset

mariathasan_dataset <- get_Mariathasan2018_PDL1_treatment()
mariathasan_dataset
```
# Session info

```{r sessionInfo}
sessionInfo()
```
