---
title: "easier data"
author: "Oscar Lapuente-Santana"
package: easierData
output: 
    rmarkdown::html_document
vignette: >
    %\VignetteIndexEntry{easier data}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load easier data

The `easierData` data package includes two main data objects:
  
* **easier_internal**: internal data required for functional performance of EaSIeR framework within the package.
* **Mariathasan2018_PDL1_treatment**: exemplary bladder cancer dataset with samples from 192 patients. The processed data is publicly available from Mariathasan et al. "TGF-B attenuates tumour response to PD-L1 blockade by contributing to exclusion of T cells", published in Nature, 2018 [doi:10.1038/nature25501](https://doi.org/10.1038/nature25501) via [`IMvigor210CoreBiologies`](http://research-pub.gene.com/IMvigor210CoreBiologies/) package under the CC-BY license.

The contents of the package can be seen by querying the annotaitonHub for the package name
```{r}
# suppressPackageStartupMessages({
#     library(ExperimentHub)
#     library(SummarizedExperiment)
#     library(easierData)
# })

# eh <- ExperimentHub()
# query(eh, "easierData")
```

The individual data objects can be accessed using either their ExperimentHub
accession number, or the convenience functions provided in this package. For
instance to access the Mariathasan2018_PDL1_treatment example dataset:

```{r}
# mariathasan_dataset <- eh[[""]]
# mariathasan_dataset
# mariathasan_dataset <- Mariathasan2018_PDL1_treatment()
# mariathasan_dataset
```
# sessionInfo()

```{r sessionInfo}
sessionInfo()
```
