---
title: "easier Data"
author:
- name: Óscar Lapuente-Santana
  affiliation: 
  - &id Computational Biology group, Department of Biomedical Engineering,
    Eindhoven University of Technology (BME, TU/e)
  email: o.lapuente.santana@tue.nl
- name: Federico Marini
  affiliation:
  - Institute of Medical Biostatistics, Epidemiology and Informatics 
    (IMBEI, Mainz)
  email: marinif@uni-mainz.de
- name: Arsenij Ustjanzew
  affiliation: 
  - Institute of Medical Biostatistics, Epidemiology and Informatics 
    (IMBEI, Mainz)
  email: arsenij.ustjanzew@uni-mainz.de
- name: Francesca Finotello
  affiliation: 
  - Institute of Bioinformatics, Biocenter Medical University of Innsbruck
  email: francesca.finotello@i-med.ac.at
- name: Federica Eduati
  affiliation: 
  - *id
  - Institute for Complex Molecular Systems, Eindhoven University of 
    Technology (ICMS, TU/e)
  email: f.eduati@tue.nl
date: "`r Sys.Date()`"
package: easierData
output: 
  pdf_document:
    latex_engine: xelatex
    toc: yes
    number_sections: true
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    theme: lumen
bibliography: references_easierData.bib
vignette: >
    %\VignetteIndexEntry{easier data}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library("easierData")
```

# Intro to `easierData`

The `easierData` package includes an exemplary cancer dataset from
@Mariathasan2018 to showcase the `easier` package:

* **Mariathasan2018_PDL1_treatment**: exemplary bladder cancer dataset 
with samples from 192 patients. This is provided as a `SummarizedExperiment` 
object containing:

  - Two assays: `counts` and `tpm` expression values.
  - Additional sample metadata in the `colData` slot, including pat_id 
    (the id of the patient in the original study), BOR, and
    TMB (Tumor Mutational Burden).
  
  The processed data is publicly available from Mariathasan et al. 
  "TGF-B attenuates tumour response to PD-L1 blockade by contributing to
  exclusion of T cells", published in Nature, 2018
  [doi:10.1038/nature25501](https://doi.org/10.1038/nature25501) via 
  [IMvigor210CoreBiologies](http://research-pub.gene.com/IMvigor210CoreBiologies/)
  package under the CC-BY license. 
  
The `easierData` data package also includes multiple data objects so-called
internal data of `easier` package since they are indispensable for the 
functional performance of the package. This includes:

* **opt_models**: the cancer-specific model feature parameters learned 
in @LAPUENTESANTANA2021100293. For each quantitative descriptor (e.g. 
pathway activity), models were trained using multi-task learning with 
randomized cross-validation repeated 100 times. For each quantitative 
descriptor, 1000 models are  available (100 per task). This is provided
as a list containing, for each cancer type and quantitative descriptor,
a matrix of feature coefficient values across different tasks.
    
* **opt_xtrain_stats**: the cancer-specific features mean and standard 
deviation of each quantitative descriptor (e.g. pathway activity) training
set used in @LAPUENTESANTANA2021100293 during randomized cross-validation 
repeated 100 times, required for normalization of the test set. This is 
provided as  a list containing, for each cancer type and quantitative 
descriptor, a matrix with feature mean and sd values across the 100 
cross-validation runs.
    
* **TCGA_mean_pancancer**: a numeric vector with the mean of the TPM 
expression of each gene across all TCGA cancer types, required for 
normalization of input TPM gene expression data.
    
* **TCGA_sd_pancancer**: a numeric vector with the standard deviation (sd)
of the TPM expression of each gene across all TCGA cancer types, required
for normalization of input TPM gene expression data.
    
* **cor_scores_genes**: a character vector with the list of genes used to 
define correlated scores of immune response. These scores were found to be
highly correlated across all 18 cancer types [@LAPUENTESANTANA2021100293].
    
* **intercell_networks**: a list with the cancer-specific intercellular 
networks, including a pan-cancer network.
    
* **lr_frequency_TCGA**: a numeric vector containing the frequency of each 
ligand-receptor pair feature across the whole TCGA database.
    
* **group_lrpairs**: a list with the information on how to group 
ligand-receptor pairs because of sharing the same gene, either as ligand
or receptor.
    
* **HGNC_annotation**: a data.frame with the gene symbols approved 
annotations obtained from https://www.genenames.org/tools/multi-symbol-checker/
[@Tweedie2021].
    
* **scores_signature_genes**: a list with the gene signatures for each score of 
immune response: CYT [@ROONEY201548], TLS [@Cabrita2020], IFNy [@Ayers2017], 
Ayers_expIS [@Ayers2017], Tcell_inflamed [@Ayers2017], Roh_IS [@Roheaah3560],
Davoli_IS [@Davolieaaf8399], chemokines [@Messina2012], IMPRES [@Auslander2018],
MSI [@Fu2019] and RIR [@JERBYARNON2018984].

# Load easier Data

Starting R, this package can be installed as follows:

```{r, eval=FALSE}
BiocManager::install("easierData")
```

The contents of the package can be seen by querying ExperimentHub for the 
package name:

```{r}
suppressPackageStartupMessages({
    library("ExperimentHub")
    library("easierData")
})

eh <- ExperimentHub()
query(eh, "easierData")
```

An overview is provided also in tabular form:

```{r}
list_easierData()
```

The individual data objects can be accessed using either their ExperimentHub
accession number, or the convenience functions provided in this package 
- both calls are equivalent. For instance to access the 
`Mariathasan2018_PDL1_treatment` example dataset:

```{r, message=FALSE}
mariathasan_dataset <- eh[["EH6677"]]
mariathasan_dataset

mariathasan_dataset <- get_Mariathasan2018_PDL1_treatment()
mariathasan_dataset
```

# Session info {-}

```{r sessionInfo}
sessionInfo()
```

# References {-}
